---
title: 'Stat 519 Homework #1 - Introduction to Multivariate Analysis, Visualization,
  Matrix Algebra'
author: "VIGNESH J MURALIDHARAN"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  word_document: default
  pdf_document: default
  html_document: default
fig_width: 6
fig_height: 3
---
#QUESTION 1
###Consider the data set LOST DAYS posted on the course homepage. 
```{r, fig.height = 3.0, fig.width = 5.0, fig.align = "center", out.width = '60%'}
LOST_DAYS =read.table('http://www.webpages.uidaho.edu/~stevel/519/Data/LOST_DAYS.txt', 
                      header=TRUE, comment.char='#')
```
###a) Examine histograms and kernel density estimates of LostDPC. In addition to the defaults, examine at least one additional choice of bin width (histogram) and bandwidth (kde). What do these plots tell you about the distribution of LostDPC?
```{r, fig.height = 3.0, fig.width = 5.0, fig.align = "center", out.width = '60%'}
attach(LOST_DAYS)
#Creating Histogram for the data Lost_Days vs Lost DPC
hist(LOST_DAYS$LostDPC,nclass=20,col='Green')
#Kernel Density Graph (Default)
plot(density(LostDPC))
#Kernel Density Graph(Bandwidth = 10)
plot(density(LostDPC, bw=10))
```

####The Kernel density estimates shows a smoother and accurate estimates of the density. In addition to the default bandwith I have addded a graph with bandwith of 10.Kernel density estimates are similar to histograms.

###b) Examine a strip chart and a boxplot of LostDPC, broken down by Power. What do these charts tell you about the effect of the use of power tools on the distribution of lost days per capita?
```{r}
#Splitting the Graph with 2 coloums 
par(mfrow=c(1,2))
#Stripchart of DPC(vs)Power
stripchart(LostDPC~Power,col='Red')
#boxplot of DPC(vc)Power
boxplot(LostDPC~Power)
```

####The Scriptchart shows that the Power"0" have more effects with the LostDPC but with Power"1" it has more effect on the LostDPC.
####The boxplot have outlieres for Power"0" having maximum data concentrated within 75 range of DPC but the Data appears to be available until 150 DPC

###c)	Examine scatterplots, looking at the effect on LostDPC (y) from each of crew size, foreman age, foreman experience, and average crew experience (x). Discuss your results. Which independent variable appears to have the greatest influence on lost days per capita?
```{r}
#Arranging the graphs (2*2) for easy visualization on the Knit file
par(mfrow=c(2,2))
plot(LostDPC~SIZE, data=LOST_DAYS, col='Red')
plot(LostDPC~ForeAge, data=LOST_DAYS, col='Blue')
plot(LostDPC~ForeExp, data=LOST_DAYS, col='Green')
plot(LostDPC~AvgExp, data=LOST_DAYS)
```

####With all the Scatter plot here the CrewSize seems to have a better influence with the LOSTDPC having a pretty nice plot having an imaginary line which can be drawn on the plot. Other plots with "ForeAge", "ForeExp", "AvgExp" have a scattered points all over the graph having not a better representations.

###d)	Use the plot3d function from the rgl library to examine 3d scatterplots of this data. Your third variable (z) should be LostDPC, and your second (y) should be the influential variable you identified in problem 3. Try each of the remaining numeric variables as your first (x) variable. (Remember, you'll need to use the rgl.snapshot command to save your rgl images to import into Word.) Discuss your results. Can you find any interaction effects (that is, behavior of x and y on z that is not obvious from just looking at x vs z and y vs z separately)?
```{r}
#For normal 3d plots images - This can be roated
library(rgl)
library(knitr)
plot3d(ForeAge,SIZE,LostDPC, type="h", size=.25, col="blue")
rgl.snapshot("D:\\University Of Idaho\\Statistical Sciences\\STAT 519 Multivariate\\plot1.png")
plot3d(ForeExp, SIZE, LostDPC, type="h", size=.25, col="red")
rgl.snapshot("D:\\University Of Idaho\\Statistical Sciences\\STAT 519 Multivariate\\plot2.png")
plot3d(AvgExp,SIZE, LostDPC, type="h", size=.25, col="green")
rgl.snapshot("D:\\University Of Idaho\\Statistical Sciences\\STAT 519 Multivariate\\plot3.png")
#For Scatter plot 3d images
library(scatterplot3d)
par(mfrow=c(2,3))
scatterplot3d(cbind(ForeAge, SIZE, LostDPC))
scatterplot3d(cbind(ForeAge, SIZE, LostDPC), type='h',
              highlight.3d=T)
scatterplot3d(cbind(ForeExp, SIZE, LostDPC))
scatterplot3d(cbind(ForeExp, SIZE, LostDPC), type='h',
              highlight.3d=T)
scatterplot3d(cbind(AvgExp,SIZE, LostDPC))
scatterplot3d(cbind(AvgExp,SIZE, LostDPC), type='h',
              highlight.3d=T)

```

####With Z-Lost_DPC we can find some interactios effects of ForeExp & Size on LostDPC. We can also slight intreactions with ForeAge. But rare amount or not much of intreactions could be found with AvgExp

###e)	Construct a pairwise scatterplot matrix, a stars plot, and a parallel coordinates plot on the five numeric variables, using color to separate by Power. Examine these plots for additional interesting interactions, outliers, or other features of the data. Discuss your observations.
```{r}
pairs(LOST_DAYS[,1:5],pch=16,col=unclass(as.factor(Power)))
stars(LOST_DAYS,key.loc=c(15,1.5), col.stars=unclass(Power)+1)
library(MASS)
parcoord(LOST_DAYS[,1:5],col=c(1,2)[unclass(as.factor(Power))])
```

####Pairs Plot - Shows all the scatter plots in one go here having the best plot with ForeAge(vs)ForeExp, This also shows the LOSTDPC Vs all the variables again having a clear idea of Size making an influencial value for it.
####Stars Plot - Shows all the observations how well each observations changes with the intreactions between them. Having more than 12 observations with almost all 5 variables having interactions. Very few observations have AvgExp as their intreation variable. This is a best graph to show and know which variable will be useful for correlation and regreassion analysis also.
####Parcord Plot - Parellel Coordinates plot produces an effective graphical summary of a multivariate data set when there are not too much variables. again all variables are connected each other. Having lot of  consistent pattern of LOSTDPC and Size.

Use non-bold regular sized fonts, insert your answer to Question 1 here ...
----------------------------------------------------------------------------------------------------------------

###End of Question 1=========================================================================================

----------------------------------------------------------------------------------------------------------------

\newpage


#QUESTION 2

###Suppose our multivariate data have sample covariance matrix 

```{r}
A = matrix(c(2,-3,2,
             -3,6,4,
             2,4,3), byrow=T,ncol=3)
A
```

###a)	Based on this covariance matrix, how many columns (variables) does the original data matrix have? Can you tell how many rows the original data matrix has?
####Answer: There are total of 3 columns the original data matrix has 3 rows basically it is a (3*3)matrix

###b)	Find the inverse of S.
```{r}
#Inverse of A
solve(A)
```
###c)	Determine the matrix $D^{-1/2}$, where $D^{-1/2}$ is defined in the text
```{r}
Dz=matrix(c(sqrt(2),0,0,0,sqrt(6),0,0,0,sqrt(3)),ncol=3)
Dz
```
###d) Find and write the correlation matrix R for this data set.
```{r}
#based on the equation R=D^-1/2 * S * D^-1/2
Rz=solve(Dz)%*%A%*%solve(Dz)
Rz
```
###End of Question 2=========================================================================================

\newpage

#QUESTION 3
###Is 
```{r}
B=matrix(c(5/13,12/13,-12/13,5/13), byrow=T,ncol=2)
```
###an orthogonal matrix or not? Check.
```{r}
t(B)
B%*%t(B)
t(B)%*%B
```
####YES IT IS IDENTICAL SO IT IS ORTHOGONAL MATRIX

###End of Question 3=========================================================================================

\newpage

#QUESTION 4

###Let
```{r}
C=matrix(c(9,-2,-2,6), byrow=T,ncol=2)
C
```
###a) Is A symmetric?
```{r}
t(C)
```
####YES IT IS SYMMETRIC C=C'
###b) Show that A is positive definite.
####With Definition Positive Definite(A>0) if t(x)*A*x>0 for all x not equal to 0
```{r}
x=matrix(c(1,2),byrow=T,ncol=2)
x
t(x)
x%*%C%*%t(x)
eigen(C)
```
####YES IT IS POSITIVE DEFINITE But, a symmetric matrix A is said to be positive definite (A>0) if all its eigen values are positive

###End of Question 4=========================================================================================

\newpage

#QUESTION 5
###Let
```{r}
D=matrix(c(4,8,8,3,6,9), byrow=T,ncol=3)
D
```
###a) Compute $BB'$ and obtain its eigenvalues and eigenvectors.
```{r}
Q5a=D%*%t(D)
Q5a
eigen(Q5a)
```
###b) Compute $B'B$ and obtain its eigenvalues and eigenvectors.
```{r}
Q5b=t(D)%*%(D)
Q5b
eigen(Q5b)
```
###c) Obtain the singular value decomposition of $B$

```{r}
svd(D)
```

###End of Question 5=========================================================================================

\newpage

#QUESTION 6
##Suppose that the determinant of a square symmetric matrix $A$ is non-zero. Does this implies that all eigenvalues of $A$ are positive?
```{r}
#Example1
Example1=matrix(c(9,-10,-10,6), byrow=T,ncol=2)
Example1
eigen(Example1)

```
####if the matrix is not symmetric then they are not real. So basically the eigen the matrix should be symmetric to be positive eigen values.
####if it is symmetric then eigen values are all real.
####EVEN IF THE DETERMINANT OF SQUARE SYMMETRIC MATRIX IS NON ZERO NOT ALL EIGEN VALUES OF A MATRIX ARE POSITIVE 

###End of Question 6=========================================================================================

\newpage

#QUESTION 7
####Consider an arbitrary $n \times p$ matrix $A$.
```{r}
#Considering the Arbitrary matrix (P*A)
E<-matrix( c( 2, -1, 0, -1, 2, -1, 0, -1, 2 ), nrow=3, byrow=TRUE )
E
```
####a) Show that $A'A$ is a symmetric $p \times p$ matrix.
```{r}
#A symmetric matrix is a square matrix that is equal to its transpose. Formally, matrix A is symmetric if E=E'
t(E)
#YES IT IS A SYMMETRIC MATRIX WHEN TRANSPOSED 
#E'*E IS A SYMMETRIC MATRIX HAVING (P*P)
F=t(E)%*%E
F
```
####AS PER THE DEF: A SQUARE MATRIX A IS CALLED SYMMETRIC IF (A = A')

###b) Show that $A'A$ is necessarily nonnegative definite.
####For a positive semi-definite matrix, the eigenvalues should be non-negative. The R function eigen is used to compute the eigenvalues. If any of the eigenvalues is less than zero, then the matrix is not positive semi-definite. Otherwise, the matrix is declared to be positive semi-definite.
#(nonnegative definite - > if and only if eighen values of A are all positive.
```{r}
#EIGEN VALUES OF THE MATRIX E CONSIDERED
eigen(E)
#EIGEN VALUES OF THE TRANSPOSED MATRIX
eigen(t(E))
#EIGEN VALUES OF THE E'E
eigen(F)
```

####THUS IT IS A POSITIVE SEMI-DEFINITE OR NON-NEGATIVE DEFINITE

###End of Question 7=========================================================================================


#The END
