---
title: "Multivariate Analysis on a Census Data of district in India"
author: "VIGNESH J MURALIDHARAN"
date: "May 7, 2018"
output:
  html_document: default
  word_document: default
---

#CONTENTS

1. INTRODUCTION
2. EXPLANATION OF VARIABLES
3. COMPARITIVE ANALYSIS USED
   A. PRINCIPLE COMPONENT ANALYSIS
   B. MULTIDIMENTIONAL SCALE ANALYSIS
4. EDUCATION - LOCATION BASED ANALYSIS
   4.1.A. EDUCATION ANALYSIS ON MOST SCATTERED AREAS IN CHI-SQUARE ANALYSIS
   4.1.B. EDUCATION ANALYSIS ON MOST SCATTERED AREAS IN PRINCIPLE COMPONENT ANALYSIS
   4.1.C. EDUCATION ANALYSIS ON MOST SCATTERED AREAS IN MULTIDIMENTIONAL SCALLING
   4.2.A. EDUCATION ANALYSIS ON CLOSELY SCATTERED AREAS IN CHI-SQUARE ANALYSIS
   4.2.B. EDUCATION ANALYSIS ON CLOSELY SCATTERED AREAS IN PRINCIPLE COMPONENT ANALYSIS
   4.2.C. EDUCATION ANALYSIS ON CLOSELY SCATTERED AREAS IN MULTIDIMENTIONAL SCALLING
5. WORKER - LOCATION BASED ANALYSIS
   5.1.A. WORKER ANALYSIS ON MOST SCATTERED AREAS IN CHI-SQUARE ANALYSIS
   5.1.B. WORKER ANALYSIS ON MOST SCATTERED AREAS IN PRINCIPLE COMPONENT ANALYSIS
   5.1.C. WORKER ANALYSIS ON MOST SCATTERED AREAS IN MULTIDIMENTIONAL SCALE ANALYSIS
   5.2.A. WORKER ANALYSIS ON CLOSELY SCATTERED AREAS IN CHI-SQUARE ANALYSIS
   5.2.B. WORKER ANALYSIS ON CLOSELY SCATTERED AREAS IN PRINCIPLE COMPONENT ANALYSIS
   5.2.C. WORKER ANALYSIS ON CLOSELY SCATTERED AREAS IN MULTIDIMENTIONAL SCALE ANALYSIS
6. PARTITION CLUSTERING 
   6.A. K-MEANS CLUSTERING FOR THE WHOLE DATASET FOR ALL LOCATIONS
   6.B. K-MEANS CLUSTERING FOR THE EDUCATION POPULATION OF THE WHOLE DATASET FOR ALL LOCATIONS
   6.C. K-MEANS CLUSTERING FOR THE WORKER POPULATION OF THE WHOLE DATASET FOR ALL LOCATIONS
7. SUMMARY
8. CONCLUSION
9. FUTURE WORK



1. INTRODUCTION

The purpose of this project is to analyze how people in India with the particular area are more related to different types of education and worker population with each city in a district 
The district name is Madras/Chennai with 7.78 million people in 270 SQ miles. Having said it is second most literate distict in the country with fifth highest employment creating district it has 35 cities as rows in the data and all the population subgroups as based on the total population in the group.

In this scenario its better to see which variable really makes cities different from each other in the particular district. Though geographical locations are not much apart, the people in each location have different sets of subgroups dependent on each other either with workculture or the education systems. All the population subgroups has been converted to proportions to make sense with the each locations.

2. EXPLANATION OF VARIABLES

   1 - Name of City block
   2 - Number of Households
   3,4,5 - Total population persons, Male & Female
   6,7,8 - Population age groups based on education (0-6),(7-13),(14-20)
   9 to 14 - Caste based (Religion based) & Tribal based education systems
   15 - 22 - Literates & Illiterate in Total with Male & Female
   23 - Total working population
   24 - 29 - Main worker
             Main Agriculture
             Main Household
             Main Cultivator
             Main Otherworking people
   30 - 34 - Marginal worker
             Marginal Agriculture
             Marginal Household
             Marginal Cultivator
             Marginal otherworking people
   35 - Non working population         
    
CASTE BASED - Though people are educated in differnt types like schools, college and medicine ..etc.. India has a special based of eduction systems for people who are interested in religious based education. For example: person who wanted to learn more about bible or quran or vedas(hindu) they have seperate education systems and they are alloted with different schools and they are considered as literate population in the whole total population subgroups.

TRIBAL BASED - These people are having tribal based education systems like forest studies, wild animal activities etc.. but since this is a city we have very less tribal population considred in this district.

Main & Marginal - Main worker are population who are working a job as there main profession 
Marginal is something like a part-time job


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

3. TOTAL COMPARITIVE ANALYSIS 
A. PRINCIPLE COMPONENT ANALYSIS 

The PCA for the total population has been done checking what variables are really making difference in the analysis the PC1 explains almost 62% of the needed analysis of the data and rest 19% from the PC2

```{r }
library(tidyverse)
library(googlesheets)

data.url = "https://docs.google.com/spreadsheets/d/1zbVfT_St4rlKpD-yXYMbcnbhqV69SJon-sGipiQoQss/edit#gid=2138857311"

#my_sheets = gs_ls()
data = data.url %>%  
  gs_url() %>% 
  gs_read()

editpurpose=data

attach(editpurpose)
row.names(editpurpose)<-editpurpose$Name
census.pca=prcomp(editpurpose[,-1],scale=TRUE)
summary(census.pca)
```

after doing PCA its better to look over in the biplot graph 
```{r}
#standard deviation of each component
census.sd=census.pca$sdev
census.var=census.pca$sdev^2
census.var
#proportion of variance explained
pve=census.var/sum(census.var)
pve
#biplot
biplot(census.pca)
#proportion of variance explained
plot(pve,xlab="principal component",ylab="Proportion of variance" 
     , ylim=c(0,1), type='b')
```

The principle component graph condenses to total of 2 PCA's and while analyzing the biplot a condensed cluster is formed making it few locations away from most of the variables with both worker and the education people. but the locations like St.Thomas mount and walajabad makes scattered locations in the biolopt which makes interest in see what really makes these locations different. 
While considering difference the most scattered locations are taken and the most clustered locations are taken to see what variables make even these locations different within each groups. 

B. MULTIDIMENTIONAL SCALLING 
Here the same analysis as PCA was checked to see if the locations are scattered in the same way PCA has analyzed the data.

```{r}
########################
#Total - multidimentinal scalling
#########################
library(MASS)
census<-as.matrix(editpurpose[,-1])
census.mds<-cmdscale(census,k=2,eig=TRUE)
census.mds$points
plot(census.mds$points,type="n",col='black')
#text(census.mds$points,rownames(census.mds$points),col=rainbow(18))
text(census.mds$points,rownames(editpurpose))

```

Here in MDS as in PCA the locations are splitted according to the same analysis and this proves that both techniqes helps us to visualize the same thing. So here the most scattered locations are taken analyzed with 2 different groups.

This census data has Education based groups and worker based groups. so, we would first pick the locations and analyzed based on the education first and then with worker in both MDS, PCA and Chi-square analysis.

4. EDUCATION - LOCATION BASED ANALYSIS

Most scattered locations 
  1. THIRVOTIUR
  2. MANALI
  3. MADAVARAM
  4. TEYNAMPET
  5. KODAMBAKAM
  6. ADAYAR
  7. POONAMALI
  8. WALAJABAD
  9. ST.THOMAS MOUNT
  10. SRIPERAMBATUR
Most clustered locations
  1. ROYAPURAM
  2. THIRUVIKANAGAR
  3. AMBATHUR
  4. ANNA NAGAR
  5. ALANDUR
  6. AYYAPANTHANGAL
  7. KELAMBAKAM
  8. KATTANGALATURE
  9. VANDALUR
  10 . TONDIRAPET

  
```{r}
#function
chisqD <- function(x) {
  r <- nrow(x)
  c<- ncol(x)
  row.sums <- apply(x,1,sum)
  col.sums <- apply(x,2,sum)
  N<-sum(row.sums)
  pijrow <- matrix(0,nrow=r,ncol=c)
  pijcol <- matrix(0,nrow=r,ncol=c)
  distx.row <- matrix(0,nrow=r,ncol=r)
  distx.col <- matrix(0,nrow=c,ncol=c)
  for (i in 1:r){
    pijrow[i,] <- x[i,]/row.sums[i] 
  }
  for (j in 1:c){
    pijcol[,j] <- x[,j]/col.sums[j] 
  }
  
  for (i in 1:r){
    for (ii in 1:(i-1)) {
      d.row<- sum( (N/col.sums)*(pijrow[i,]-pijrow[ii,])^2 ) 
      distx.row[i,ii] <- d.row
      distx.row[ii,i] <- d.row
    }
  }
  
  for (j in 1:c){
    for (jj in 1:(j-1)) {
      d.col <- sum( (N/row.sums)*(pijcol[,j]-pijcol[,jj])^2 ) 
      distx.col[j,jj] <- d.col
      distx.col[jj,j] <- d.col
    }
  }
  return((list(dist.row=distx.row, dist.col=distx.col))) 
}

```

4.1.A. EDUCATION ANALYSIS ON MOST SCATTERED AREAS IN CHI-SQUARE ANALYSIS 

In the most scattered areas the tribal education system is very less which makes sense for sure because its a city based census. But comparitively the literates population is very high and at the same time non working population is also very high. Unfortunatelly though the illiterate population is very less the number of working populatin is very high is understable because even if people are not studying they can work something in their life. Here in India caste based education is considered as the population who are literates so these kind of population is higher in there areas . 

```{r}
data.url = "https://docs.google.com/spreadsheets/d/1mF1QWTnOEmMugQ_u30rY-4AAFkoLOgGRbie1AeHWHwI/edit#gid=920190064"

#my_sheets = gs_ls()
data = data.url %>%  
  gs_url() %>% 
  gs_read()

chisquare1to9=data
chisquare1to9

attach(chisquare1to9)
dim(chisquare1to9)
cities<-as.matrix(chisquare1to9[,c(2:10)])
t(cities)
G<-t(cities)%*%cities
chisqD(G)$dist.col
rG1 <- cmdscale(chisqD(G)$dist.col, eig = TRUE)
cG1 <- cmdscale(chisqD(G)$dist.row, eig = TRUE)
plot(rG1$points, xlim = range(rG1$points[,1], cG1$points[,1]) , 
     ylim = range(rG1$points[,1], cG1$points[,1]))
text(rG1$points, labels = colnames(G), cex = 1)
abline(h = 0, lty = 2,col='red')
abline(v = 0, lty = 2,col='red')
```

4.1.B. EDUCATION ANALYSIS ON MOST SCATTERED AREAS IN PRINCIPLE COMPONENT ANALYSIS

In the PCA here we can clearly see that chi-square method proves us the same for the literates population is higher in areas like st.Thomas mount , Teynampet, kodambakam and Adyar. rather than the caste based education is seems higher in poonamali and population age group of (0to6) is higher in madavaram while thirvotur seems like having all the variables in common.

```{r}
row.names(chisquare1to9)<-chisquare1to9$Name
census1to9.pca=prcomp(chisquare1to9[,-1],scale=TRUE)
summary(census1to9.pca)
#Standard deviation of each component
census1to9.sd=census1to9.pca$sdev
census1to9.var=census1to9.pca$sdev^2
census1to9.var
#proportion of variance explained
pve=census1to9.var/sum(census1to9.var)
#biplot
biplot(census1to9.pca)
#proportion of variance explained
plot(pve,xlab="principal component",ylab="Proportion of variance" , ylim=c(0,1), type='b')
```

4.1.C. EDUCATION ANALYSIS ON MOST SCATTERED AREAS IN MULTIDIMENTIONAL SCALLING

Areas in MDS proves the same scattered locations in PCA has the same adjustments in the MDS scale. While manali (2) is totally different with respect to total worker population but in MDS scale it seems to share a bit with other nearby locations like teynampet(4) and kodambakam(5).

```{r}
census1<-as.matrix(chisquare1to9[,-1])
census1.mds<-cmdscale(census1,k=2,eig=TRUE)
census1.mds$points
row.names(chisquare1to9)<-chisquare1to9$Name
plot(census1.mds$points,type="n",col='red')
#text(census.mds$points,rownames(census.mds$points),col=rainbow(18))
text(census1.mds$points,rownames(chisquare1to9))
```

4.2.A  EDUCATION ANALYSIS ON CLOSELY SCATTERED AREAS IN CHI-SQUARE ANALYSIS

Here the caste based education is most higher compared to the normal literate population in the cities. Which means that caste based population is also more here who are insisting the yonger generations on these types of education systems in the society. At the same time you can also see the comparison of the population education age group of (7-13) is very less in the total population which clearly shows that entry level school education is completely not insisted in the society here. Otherwise all other population groups occur in similar way for all locations.

```{r}
data.url = "https://docs.google.com/spreadsheets/d/1awOEtfoFg7r7tVMx-gc5G5xez1z107gvfzG1t7bY-t8/edit#gid=1676744121"

#my_sheets = gs_ls()
data = data.url %>%  
  gs_url() %>% 
  gs_read()

chisquaresecond9=data
chisquaresecond9

attach(chisquaresecond9)
dim(chisquaresecond9)
cities1<-as.matrix(chisquaresecond9[,c(2:10)])
dim(cities1)
t(cities1)
G1<-t(cities1)%*%cities1
G1
chisqD(G1)$dist.col
rG2 <- cmdscale(chisqD(G1)$dist.col, eig = TRUE)
cG2 <- cmdscale(chisqD(G1)$dist.row, eig = TRUE)
plot(rG2$points, xlim = range(rG2$points[,1], cG2$points[,1]) , 
     ylim = range(rG2$points[,1], cG2$points[,1]))
text(rG2$points, labels = colnames(G1), cex = 1)
abline(h = 0, lty = 2,col='red')
abline(v = 0, lty = 2,col='red')
```

4.2.B. EDUCATION ANALYSIS ON CLOSELY SCATTERED AREAS IN PRINCIPLE COMPONENT ANALYSIS

Here we can clearly see that the illitrate population is drastically gone down in towns like Ayyapandhangal(6) & Kattangalatur(8). But very high literate population in Annanagar(4), Alandur(5), vandalur(9). With all this even interesting things like population age group of education for (14-20) is also more in Royapuram. 
Which means compared to Royapuram(1) & Ayyapandhangal (6) we should have more elder people who are illiterate but yonger generations are giving importance to education.

```{r}
row.names(chisquaresecond9)<-chisquaresecond9$Name
censussecond9.pca=prcomp(chisquaresecond9[,-1],scale=TRUE)
summary(censussecond9.pca)
print(censussecond9.pca)
#Standard deviation of each component
censussecond9.sd=censussecond9.pca$sdev
censussecond9.var=censussecond9.pca$sdev^2
censussecond9.var
#proportion of variance explained
pve=censussecond9.var/sum(censussecond9.var)
pve
biplot(censussecond9.pca) #biplot
#proportion of variance explained
plot(pve,xlab="principal component",ylab="Proportion of variance" , ylim=c(0,1), type='b')
```

4.2.C. EDUCATION ANALYSIS ON CLOSELY SCATTERED AREAS IN MULTIDIMENTIONAL SCALLING

Here in MDS we have a good comparison scatter with regard to PCA also having kelambakam(7) spotted all alone. And all the points taken into analysis have a similar effect which was shown in the PCA analysis. So this proves exactly better way of my work.

```{r}
census2<-as.matrix(chisquaresecond9[,-1])
census2.mds<-cmdscale(census2,k=2,eig=TRUE)
census2.mds$points
row.names(chisquaresecond9)<-chisquaresecond9$Name
plot(census2.mds$points,type="n",col='red')
#text(census.mds$points,rownames(census.mds$points),col=rainbow(18))
text(census2.mds$points,rownames(chisquaresecond9))
```

5. WORKER - LOCATION BASED ANALYSIS 

5.1.A. WORKER ANALYSIS ON MOST SCATTERED AREAS IN CHI-SQUARE ANALYSIS 

With otherworking population more here and agriculture population is very less we can clearly see that when comparing with the education of most scattered the literate population is more so the other working population and worker and household population is more. The cultivation is suprizingly getting almost touched the margin which means that in these areas the cultivators are having agricultural lands in these areas. In a city this is very rare to see. 

```{r}
data.url = "https://docs.google.com/spreadsheets/d/1EfJamAr67K1KZun8wzYAwVGxjtmWZZYSUqPgIXRz3c0/edit#gid=1094962871"

#my_sheets = gs_ls()
data = data.url %>%  
  gs_url() %>% 
  gs_read()
chisquareworker1to9=data
chisquareworker1to9

attach(chisquareworker1to9)
dim(chisquareworker1to9)
worker1<-as.matrix(chisquareworker1to9[,c(2:11)])
dim(worker1)
t(worker1)
G2<-t(worker1)%*%worker1
G2
chisqD(G2)$dist.col
rG3 <- cmdscale(chisqD(G2)$dist.col, eig = TRUE)
cG3 <- cmdscale(chisqD(G2)$dist.row, eig = TRUE)
plot(rG3$points, xlim = range(rG3$points[,1], cG3$points[,1]) , 
     ylim = range(rG3$points[,1], cG3$points[,1]))
text(rG3$points, labels = colnames(G2), cex = 1)
abline(h = 0, lty = 2,col='red')
abline(v = 0, lty = 2,col='red')
```

5.1.B. WORKER ANALYSIS ON MOST SCATTERED AREAS IN PRINCIPLE COMPONENT ANALYSIS 

Here in this PCA most of the population is based on the otherworking groups than rest over subgroups. Only St Thomas mount(8) have a main worker based population and Sripermbadur(9) have marginal other worker population. Other distribution of the locations i think its based on the the divitions of the main working groups. In this scenario Though we have more literate population in the most scattered areas the people who work for other working is more which means on aveage it can be said that competitions in getting a dream job is very difficult though being a literate. Or this can also be because of the illiterate pop who needs other working income might be the way this population is divided. In either case its really a very diverse not census group.

```{r}
row.names(chisquareworker1to9)<-chisquareworker1to9$Name
workerfirst9.pca=prcomp(chisquareworker1to9[,-1],scale=TRUE)
summary(workerfirst9.pca)
print(workerfirst9.pca)
#Standard deviation of each component
workerfirst9.sd=workerfirst9.pca$sdev
workerfirst9.var=workerfirst9.pca$sdev^2
workerfirst9.var
#proportion of variance explained
pve=workerfirst9.var/sum(workerfirst9.var)
pve
#biplot
biplot(workerfirst9.pca)
#proportion of variance explained
plot(pve,xlab="principal component",ylab="Proportion of variance" , ylim=c(0,1), type='b')
```

5.1.C. WORKER ANALYSIS ON MOST SCATTERED AREAS IN MULTIDIMENTIONAL SCALE ANALYSIS 

As explained in the PCA though the group are similar in few things in MDS it is totally differnt story of showing various points for the locations. But still it proves a little bit for few locations compared to the PCA. 

```{r}
worker2<-as.matrix(chisquareworker1to9[,-1])
worker2.mds<-cmdscale(worker2,k=2,eig=TRUE)
worker2.mds$points
row.names(chisquareworker1to9)<-chisquareworker1to9$Name
plot(worker2.mds$points,type="n",col='red')
#text(census.mds$points,rownames(census.mds$points),col=rainbow(18))
text(worker2.mds$points,rownames(chisquareworker1to9))
```

5.2.A. WORKER ANALYSIS ON CLOSELY SCATTERED AREAS IN CHI-SQUARE ANALYSIS 

Intrestingly, marginal agri workers are more in this population groups. that's would be a real suprize because, when we compare with the PCA of the education of closely scatterd group I mentioned that elder people were not much educated but only the yonger. So in this scenario the elder people are considered doing agriculture and cultivators and only the yonger generations are having the household and other working interest.

```{r}
data.url = "https://docs.google.com/spreadsheets/d/1581Srpvo9_TSFPr6Oc5sWtYQD4eUav1rJCTEsyzxSH8/edit#gid=2041523407"

#my_sheets = gs_ls()
data = data.url %>%  
  gs_url() %>% 
  gs_read()
chisquareworkersecond9=data
chisquareworkersecond9

attach(chisquareworkersecond9)
dim(chisquareworkersecond9)
worker3<-as.matrix(chisquareworkersecond9[,c(2:11)])
dim(worker3)
t(worker3)
G3<-t(worker3)%*%worker3
G3
chisqD(G3)$dist.col
rG4 <- cmdscale(chisqD(G3)$dist.col, eig = TRUE)
cG4 <- cmdscale(chisqD(G3)$dist.row, eig = TRUE)
plot(rG4$points, xlim = range(rG4$points[,1], cG4$points[,1]) , 
     ylim = range(rG4$points[,1], cG4$points[,1]))
text(rG4$points, labels = colnames(G3), cex = 1)
abline(h = 0, lty = 2,col='red')
abline(v = 0, lty = 2,col='red')
```

5.2.B. WORKER ANALYSIS ON CLOSELY SCATTERED AREAS IN PRINCIPLE COMPONENT ANALYSIS 

Tondiapet(1) has shown very less main worker population which makes it way different from the other cities. But otherwise the central cluster cities showing very les of other working groups and making all the Kelambakam (7), Ayyapandangal (6) as the working population group. which is very good to see with respect to the eduction analysis comparison having illiterate population. So though people are illiterte the working population is always higher. 

```{r}
row.names(chisquareworkersecond9)<-chisquareworkersecond9$Name
workersecond10.pca=prcomp(chisquareworkersecond9[,-1],scale=TRUE)
summary(workersecond10.pca)
print(workersecond10.pca)
#Standard deviation of each component
workersecond10.sd=workersecond10.pca$sdev
workersecond10.var=workersecond10.pca$sdev^2
workersecond10.var
#proportion of variance explained
pve=workersecond10.var/sum(workersecond10.var)
pve
#biplot
biplot(workersecond10.pca)
#proportion of variance explained
plot(pve,xlab="principal component",ylab="Proportion of variance" , ylim=c(0,1), type='b')
```

5.2.C. WORKER ANALYSIS ON CLOSELY SCATTERED AREAS IN MULTIDIMENTIONAL SCALE ANALYSIS

The analysis in the PCA with all he scattered locations exactly matches in the MDS except for the Vandalur(9) which shows all the marginal and the main cultivator making it differnt from the other locations.

```{r}
worker3<-as.matrix(chisquareworkersecond9[,-1])
worker3.mds<-cmdscale(worker3,k=2,eig=TRUE)
worker3.mds$points
row.names(chisquareworkersecond9)<-chisquareworkersecond9$Name
plot(worker3.mds$points,type="n",col='red')
#text(census.mds$points,rownames(census.mds$points),col=rainbow(18))
text(worker3.mds$points,rownames(chisquareworkersecond9))
```

Since the locations are not based on the Regions I did'nt use the Hirachial Clustering. And moreover since its a one part of location based in few distance partition clustering seems like a best option for this data analysis.

6. PARTITION CLUSTERING 

Having K-means is the simplest learning algorithms that will solve the clustering problem. The main features of k-means which make it efficient are often regarded as its Euclidean distance is used as a metric and variance as a measure of cluster scatter. The number of K is an input parameter and thats the reason its important to run diagnostic checks for determining he number of clusters in the dataset.Since it is a cluster model, the concept is based on spherical cluster center. 

6.A. K-MEANS CLUSTERING FOR THE WHOLE DATASET FOR ALL LOCATIONS.

partition clustering On the whole dataset for k-means we have 2 clusters with Dim 1 having 62% and Dim 2 having 19% 
the two clusters fromed clearly shows that the closely related locations are grouped here on the right with blue shade and yellow shade having almost similar locations available in the PCA for making as clusters. K=2 seems like a best option from the silhouettte test. 
When we see the Distance graph also we can clearly see that similar locations are close to one another. blue color corresponds to small distance and red color indicates big distance between observation for the total census dataset.Also thsi plot helps us to clearly see also the difference between the cluster locations clearly mentions here as the smaller cluster (blue) is in the blue color of smaller distance in the 4th graph.

```{r}
library(mclust)
library(MASS)
library(factoextra)
library(ggdendro)
attach(editpurpose)
library(dendextend)
censusdata.s2<-scale(editpurpose[,c(-1)])
set.seed(123)
fviz_nbclust(censusdata.s2,kmeans,method="silhouette")
fviz_nbclust(censusdata.s2,kmeans,method="wss")
censusdata.k42<-kmeans(censusdata.s2,centers=2,iter.max = 100,nstart = 25)
fviz_cluster(censusdata.k42,data=censusdata.s2,
             ellipse.type="convex",palette="jco",repel=TRUE,
             ggtheme=theme_minimal())
distance <- get_dist(censusdata.s2)
fviz_dist(distance, gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07"))

```

6.B. K-MEANS CLUSTERING FOR THE EDUCATION POPULATION OF THE WHOLE DATASET FOR ALL LOCATIONS.

Partition clustering On the education population for k-means we have 5 clusters with Dim 1 having 82% and Dim 2 having 12% 
the five clusters fromed clearly shows different proves for us that the closely related locations are grouped here on the right with blue shade and yellow shade having almost similar locations available in the PCA for making as clusters. K=2 seems like a best option from the silhouettte test.
    1. Considering number 16 in the K-means which as formed a seperate cluster if we compare with the total PCA earlier it is the only place which has the more "Tribal education" in the all locations making it as a seperate cluster 
    2. Points 14,27 and 15 in K-means which has seperate cluster when compared with the total PCA earlier there are the locations with highest illitrate population has been seen.
    3. Points 1,6,10,35 in k-means which has seperte cluster when compared with the total PCA we can see that the total caste based eduction is higher in these areas making it as a seperate cluster. 
    4. Most of the clusters seperation has been proved in the distance graph also with blue being the smaller distance and from 35 poin the the graph it is of longer distance.
    
```{r}
censusdata.s1<-scale(editpurpose[,c(6,7,8,10,13,15,19,23,35)])
set.seed(123)
fviz_nbclust(censusdata.s1,kmeans,method="silhouette")
fviz_nbclust(censusdata.s1,kmeans,method="wss")
censusdata.k41<-kmeans(censusdata.s1,centers=5,iter.max = 100,nstart = 25)
fviz_cluster(censusdata.k41,data=censusdata.s1,
             ellipse.type="convex",palette="jco",repel=TRUE,
             ggtheme=theme_minimal())
distance <- get_dist(censusdata.s1)
fviz_dist(distance, gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07"))

```

6.C. K-MEANS CLUSTERING FOR THE WORKER POPULATION OF THE WHOLE DATASET FOR ALL LOCATIONS.

Partition clustering On the worker population for k-means we have 2 clusters.  clusters fromed clearly shows different proves for us that the scattered related locations are grouped here on the right with yellow shade and blue shade having almost similar locations available in the PCA for making as clusters. K=2 seems like a best option from the silhouettte test.
    1. Considering number 14(Walajabad) &15(Sriprampudrur) was barely about to join the right(yellow) cluster when compared to the total k-means partition. This might be because these both locations have wide range of difference in the PCA earlier and that makes it close during the cluster part. 

```{r}
censusdata.s<-scale(editpurpose[,c(25,34)])
set.seed(123)
fviz_nbclust(censusdata.s,kmeans,method="silhouette")
fviz_nbclust(censusdata.s,kmeans,method="wss")
censusdata.k4<-kmeans(censusdata.s,centers=2,iter.max = 100,nstart = 25)
fviz_cluster(censusdata.k4,data=censusdata.s,
             ellipse.type="convex",palette="jco",repel=TRUE,
             ggtheme=theme_minimal())
distance <- get_dist(censusdata.s)
fviz_dist(distance, gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07"))

```

7. SUMMARY

In this dataset for census the Y-parameter has taken as only one variable that is the Total Population Persons in the district. Which means analyzing with MANOVA is not possible. But in this scenario we can either conduct the ANOVA or regression analysis. Since all the locations is based on population parameters I have considered to convert all the subgroups adding to to the total population parameter and converted into the proportions. 

So ANOVA is done with the parameters for all the population first and seen for the significant variables affecting the total population difference in the district
Those are 
  1. Y2 - Number of Households
  2. Y4 - Total Population Male
  3. Y5 - Total Population Female
  4. Y6 - Population education age group (0-6)
  5. Y7 - Population education age group (7-13)
  6. Y9 - Caste based People
  7. Y13 - Caste based education for the people
  8. Y15 - Literate Population
  9. Y28 - Main Household industry working groups
  10. Y32 - Marginal Agriculture working groups

```{r}
data.url = "https://docs.google.com/spreadsheets/d/16Ju0LWxCntYuFHMFHB4d7fknE-I6gEi2-EY02Cm7PXs/edit#gid=1139834335"

#my_sheets = gs_ls()
data = data.url %>%  
  gs_url() %>% 
  gs_read()
editpurposeys=data
editpurposeys
attach(editpurposeys)
fit<-aov(y3~y2+y4+y5+y6+y7+y8+y9+y10+y11+y12+y13+y14+y15+y16+y17+y18+y19+y20
         +y21+y22+y23+y24+y25+y26+y27+y28+y29+y30+y31+y32+y33+y34+y35+y36,data=editpurposeys)
summary(fit)
```

After analyzing this I felts it will be better to analyze which varibles really effect the eduction groups in these locations. Here all the variables with regard to eduction was significant.

```{r}
attach(editpurposeys)
fitedu<-aov(y3~y6+y7+y8+y10+y11+y15+y19+y23+y35,data=editpurposeys)
summary(fitedu)
```

Comparitive for the education its better to analyze the working variables also with all these locations. Here only Main worker(y25), Main household(y28) & Marginal agri (y32) makes significant difference within these population locations of this district.

```{r}
attach(editpurposeys)
fitwork<-aov(y3~y25+y26+y27+y28+y29+y30+y31+y32+y33+y34,data=editpurposeys)
summary(fitwork)
```

8. CONCLUSION

In conclusion I would say that though the city blocks are scattered in a near distance here geographical location doesn't seem to like have an effect in the employment nor the education for this population group. It is certain categorized people in each areas making the difference. Categorized people may be based on Poor or language based or religion based or also work based. So in total compared to real statistics in the government websites we can see tha only normal census in india doen't really help in analyzing the whole population based on the employment and education. There is some other significant variables which is affecting these groups making it different. This could be a vital analyzis between the group of this location. I have not done the hiracheal clustering because if its divided into regions i would have done it. I have not done MANOVA instead ANOVA because there is only one response variable total population. The general understanding of the data graphs can be done here,but in these kind of data creating Pairs graph with more variables took lot of time to export the image out so having more in these cases that was avoided to jump directly into the analysis.

9. FUTURE WORK

In future analyzing between the two disticts with each of same locations and groups will be a best way to use these multivariate analysis which helps us to analyze different population group of each district. I will make sure this will be done very soon and try to keep these techniques to find differences. 

#THE END

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
